ProSolo: Accurate Variant Calling from Single Cell DNA Sequencing Data

### 2021-02-05 manuscript/supplement.tex
add FACS sorting strategy figure to supplement

\begin{figure}[!tpb]
 \ContinuedFloat
 \begin{minipage}{\linewidth}
  \includegraphics[width=\linewidth]{figs/Wang2014/Wang2014_coverage_dist.pdf}
 \end{minipage}
 \caption{
 Sequencing coverage of targeted genomic sites for multiple displacement amplified (MDA) single cells and respective bulk samples (continued from previous page).
 Panel \textbf{C} shows the coverage of the single cells (a1 to a8, h1 to h8, and TNBC-n1 to TNBC-n16), and the bulk samples (TNBC-Pop-Tumor and TNBC-Pop-Normal) that were used as ground truth for the normal single cells (with validated somatic variants added) and as ground truth for the tumor single cells (with validated somatic variants removed), respectively \cite{wang_clonal_2014}.
 }
 \label{fig:coverage-dists}
\end{figure}

\textbf{Whole exome sequencing of five human granulocytes with a pedigree ground truth.}
For the second benchmarking dataset, Blood was taken from a patient after informed consent and granulocytes were sorted out in several steps: Initially, CD3- cells were captured via Magnetic-Activated Cell Sorting (MACS); among them, CD66b+ granulocytes were filtered out via Fluorescence Activated Cell Sorting (FACS) (Figure~\ref{fig:datasets}B).
Single cells from that cell population were isolated using a Fluidigm C1 microfluidics chip and subjected to multiple displacement amplification (MDA) using the GE Healthcare GenomiPhi Kit in its version adapted to that microfluidic chip.
With a panel of 16 custom primer pairs covering different genes across chromosomes used in a quantitative real-time PCR, we selected granulocytes where at least 15 of these loci were properly amplified.
For those cells, we performed whole exome capture using the Agilent SureSelectXT Human All Exon V5+UTRs Kit.
The paired-end sequencing library (2x101bp) was then prepared with the TruSeq V3 Kit and sequenced on a HiSeq 2500.
From the remaining sorted cell population, we extracted bulk DNA and submitted it to the same whole exome capture and DNA sequencing without MDA.\\

@@

\begin{center}
 \begin{table}[tbp]
  \caption{Software and related resources used in this manuscript. All software versions as installed with conda.}
  \label{tab:software-overview}
  \begin{tabular}{llll}
    \toprule
    software / resource & version   & purpose                       & citation \\
    \toprule
    cutadapt            & 1.14      & adapter trimming              & \citep{martin_cutadapt_2011} \\
    \midrule
    trimmomatic         & 0.36      & quality trimming              & \citep{bolger_trimmomatic_2014} \\
    \midrule
    bwa mem             & 0.7.15    & read alignment / mapping      & \citep{li_aligning_2013} \\
    \midrule
    samtools            & 1.4.1     & handle SAM / BAM files        & \citep{li_sequence_2009} \\
    \midrule
    picardtools         & 2.9.2     & mark duplicates               &  \\
    \midrule
    bcftools            & 1.6       & handle VCF / BCF files        & \citep{danecek_variant_2011} \\
                        & 1.10      & some rule-specific environments & \citep{danecek_variant_2011} \\
    \midrule
    mosdepth            & 0.2.5     & coverage calculations         & \citep{pedersen_mosdepth_2018} \\
    \midrule
    Okabe Ito palette   &           & colorblind-safe               & \citep{okabe_color_2008} \\
    \midrule
    ggthemes            & 0.2.5     & Okabe Ito palette             & \citep{arnold_ggthemes_2019} \\
    \midrule
    Brewer palettes     &           & colorblind-friendly           & \citep{brewer_colorbrewer_2006} \\
    \midrule
    RColorBrewer        & 1.1\_2     & Brewer palettes              & \citep{neuwirth_rcolorbrewer_2014} \\
    \midrule
    tidyverse           & 1.1.1     & ground truth comparisons      & \citep{wickham_welcome_2019} \\
                        & 1.3.0     & results data handling         &  \\
    \midrule
    ggplot2             & 3.3.0     & result plots                  & \citep{wickham_ggplot2_2016} \\
    \midrule
    Inkscape            & 0.92.4    & info-graphics, improved labels & \\
    \midrule
    VGAM                & 1.1\_1     & distributions in R            & \citep{yee_vgam_2008,yee_vector_2015} \\
    \midrule
    UpsetR              &           & set intersection plot \ref{fig:granulocytes-ground-truth} & \citep{conway_upsetr_2017} \\
    \bottomrule
  \end{tabular}
 \end{table}
\end{center}

>>>>

\begin{figure}[!tpb]
 \ContinuedFloat
 \begin{minipage}{\linewidth}
  \includegraphics[width=\linewidth]{figs/Wang2014/Wang2014_coverage_dist.pdf}
 \end{minipage}
 \caption{
 Sequencing coverage of targeted genomic sites for multiple displacement amplified (MDA) single cells and respective bulk samples (continued from previous page).
 Panel \textbf{C} shows the coverage of the single cells (a1 to a8, h1 to h8, and TNBC-n1 to TNBC-n16), and the bulk samples (TNBC-Pop-Tumor and TNBC-Pop-Normal) that were used as ground truth for the normal single cells (with validated somatic variants added) and as ground truth for the tumor single cells (with validated somatic variants removed), respectively \cite{wang_clonal_2014}.
 }
\end{figure}

\textbf{Whole exome sequencing of five human granulocytes with a pedigree ground truth.}
For the second benchmarking dataset, Blood was taken from a patient after informed consent and granulocytes were sorted out in several steps: Initially, CD3- cells were captured via Magnetic-Activated Cell Sorting (MACS); among them, CD66b+ granulocytes were filtered out via Fluorescence Activated Cell Sorting (FACS) (Figures~\ref{fig:datasets}B and \ref{fig:FACS-scheme}).
Single cells from that cell population were isolated using a Fluidigm C1 microfluidics chip and subjected to multiple displacement amplification (MDA) using the GE Healthcare GenomiPhi Kit in its version adapted to that microfluidic chip.
With a panel of 16 custom primer pairs covering different genes across chromosomes used in a quantitative real-time PCR, we selected granulocytes where at least 15 of these loci were properly amplified.
For those cells, we performed whole exome capture using the Agilent SureSelectXT Human All Exon V5+UTRs Kit.
The paired-end sequencing library (2x101bp) was then prepared with the TruSeq V3 Kit and sequenced on a HiSeq 2500.
From the remaining sorted cell population, we extracted bulk DNA and submitted it to the same whole exome capture and DNA sequencing without MDA.\\

\begin{figure}[!tpb]
 \includegraphics[width=\linewidth]{figs/Laehnemann2017/Laehnemann2017_Granulozyten_P.pdf}
 \caption{
 59 CD66b+ granulocytes were sorted out from a total population of 35779 CD3-T cell depleted peripheral blood cells (gate C in the right-most panel) and from these, five single cells were chosen where whole genome amplification had yielded DNA for at least 15 out of 16 control loci.
 }
 \label{fig:FACS-scheme}
\end{figure}

@@

\begin{center}
 \begin{table}[tbp]
  \caption{Software and related resources used in this manuscript. All software versions as installed with conda.}
  \label{tab:software-overview}
  \begin{tabular}{llll}
    \toprule
    software / resource & version   & purpose                       & citation \\
    \toprule
    cutadapt            & 1.14      & adapter trimming              & \citep{martin_cutadapt_2011} \\
    \midrule
    trimmomatic         & 0.36      & quality trimming              & \citep{bolger_trimmomatic_2014} \\
    \midrule
    bwa mem             & 0.7.15    & read alignment / mapping      & \citep{li_aligning_2013} \\
    \midrule
    samtools            & 1.4.1     & handle SAM / BAM files        & \citep{li_sequence_2009} \\
    \midrule
    picardtools         & 2.9.2     & mark duplicates               &  \\
    \midrule
    bcftools            & 1.6       & handle VCF / BCF files        & \citep{danecek_variant_2011} \\
                        & 1.10      & rule-specific environment     &  \\
    \midrule
    mosdepth            & 0.2.5     & coverage calculations         & \citep{pedersen_mosdepth_2018} \\
    \midrule
    Okabe Ito palette   &           & colorblind-safe               & \citep{okabe_color_2008} \\
    \midrule
    ggthemes            & 0.2.5     & Okabe Ito palette             & \citep{arnold_ggthemes_2019} \\
    \midrule
    Brewer palettes     &           & colorblind-friendly           & \citep{brewer_colorbrewer_2006} \\
    \midrule
    RColorBrewer        & 1.1\_2     & Brewer palettes              & \citep{neuwirth_rcolorbrewer_2014} \\
    \midrule
    tidyverse           & 1.1.1     & ground truth comparisons      & \citep{wickham_welcome_2019} \\
                        & 1.3.0     & results data handling         &  \\
    \midrule
    ggplot2             & 3.3.0     & result plots                  & \citep{wickham_ggplot2_2016} \\
    \midrule
    Inkscape            & 0.92.4    & info-graphics, improved labels & \\
    \midrule
    VGAM                & 1.1\_1     & distributions in R            & \citep{yee_vgam_2008,yee_vector_2015} \\
    \midrule
    UpsetR              &           & set intersection plot \ref{fig:granulocytes-ground-truth} & \citep{conway_upsetr_2017} \\
    \bottomrule
  \end{tabular}
 \end{table}
\end{center}
